---
title: "R for Humans, Ensemble Modeling, and Deep learning"
author: " "
date: "February 15th, 2018"
output:
  ioslides_presentation:
    widescreen: true
    logo: images/Rlogo.png

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```


## So what can R do? {.columns-2 .smaller}

**_Blind Men Appraising an Elephant_**, by Ohara Donshu

Like _Blind Men Appraising an Elephant_, its natural for experts spanning a diverse range of disciplines to view Data Science from the perspective of their own domain.

However, like Ohara Donshu, those interested in Data Science needs to be able to view the subject from a far, and understand many parts make a whole.



```{r, out.width = "550px"}
path <- paste0(getwd(), "/images/Blind_Men_Appraising.jpg")
knitr::include_graphics(path)
```

[Image Source: wiki/Blind_men_and_an_elephant](https://en.wikipedia.org/wiki/Blind_men_and_an_elephant)



## Tonights talks will show us

**Best practices in writing readable & maintainable R code - Peter Hurford**


**Ensemble modeling, sports analytics & hedge funds - Igor Veksler**

**Deep Learning with R using Keras - Rajiv Shah**




# Some Brief History


## 1880: Statistician solves a Big Data problem {.columns-2 .smaller}

**1880** Since its inception in 1790, the U.S. census has expanded dramatically. It takes 8 years to process.

Herman Hollerith, **_a Statistician of the 1880 census_**, sets out to process it faster. Inspired by railway conductors, he creates punch cards to collect data. Inserted into a machine, the holes in each card allow an electrical circuit to complete, advancing a counting wheel for each topic and card.

**1890** The first census to use Hollerith's Tabulating Machine is a success, completing in just 1 year. 

**1896** Hollerith forms the "Tabulating Machine Company", which later becomes "International Business Machines" better known today as "IBM".


```{r, out.width = "400px"}
path <- paste0(getwd(), "/images/HollerithMachine.jpg")
knitr::include_graphics(path)
```

Sources: [Census.gov/history](https://www.census.gov/history/www/census_then_now/notable_alumni/herman_hollerith.html) and pg. 36 _The Innovators_, Walter Isaacson



## Ensemble Methods {.columns-2 .smaller}

Leo Breiman, Jerome H. Friedman, and others bridge the gap between statistics and computer science with Classification and Regression Trees.

**1984**: Classification and Regression Trees, Leo Breiman and Jerome H. Friedman

**1994**: Bagging Predictors, Leo Breiman

**1999**: Random Decisions Forests

**2000**: Additive Logistic Regression: A Statistical View of Boosting - Friedman, Jerome, Hastie, Trevor, Tibshirani, Robert 

**2001**: Statistical Modeling: _The Two Cultures_. Breiman.

[Source Revolution Analytics blog: A Thumbnail History of Ensemble Methods](http://blog.revolutionanalytics.com/2014/03/a-thumbnail-history-of-ensemble-methods.html)

```{r, out.width = "300px"}
path <- paste0(getwd(), "/images/CART-Book.jpg")
knitr::include_graphics(path)
```



## John Tukey, a Data Science pioneer {.columns-2 .smaller}

**1962** [_The Future of Data Analysis_](https://www.jstor.org/stable/2237638) 

**_I have come to feel that my central interest is in data analysis, which I take to include, among other things: procedures for analyzing data, techniques for interpreting the results of such procedures, ways of planning the gathering of data to make its analysis easier, more precise or more accurate, and all the machinery and results of (mathematical) statistics which apply to analyzing data._**

**1964** [_The technical tools of statistics_](http://ect.bell-labs.com/sl/tukey/memo/techtools.html) 

**_Today, software and hardware together provide far more powerful factories than most statisticians realize, factories that many of today's most able young people find exciting and worth learning about on their own._**



**1939** Earns a Ph.D Mathematics, Princeton

**1947** Working with [Claude Shannon](https://en.wikipedia.org/wiki/Claude_Shannon) at **AT&T Bell Laboratories**, Tukey Invents the term "bit"

**1956**  Sets up Princeton's Statistical Techniques Research Group (STRG)

**1961** Promoted to Executive Director-Research Information Sciences at **AT&T Bell Laboratories**

**1965** 1st Chair of Princeton's Statistics Department

**1965** Published influential paper on the Fast Fourier Transformation with J.W. Cooley, leading to the rapid development of digital signal processing

[Source: Memories of John W. Tukey](http://ect.bell-labs.com/sl/tukey/)


## 1977: John Tukey's Exploratory Data Analysis {.columns-2 .smaller}

**"Since the aim of exploratory data analysis is to learn what seems to be, it should be no surprise that pictures play a vital role in doing it well. There is nothing better for making you think of questions you had forgotten to ask (even mentally)"** - John Tukey

Tukey's championing of EDA encouraged the development of statistical computing packages, especially S at Bell Labs. The S programming language inspired the systems 'S'-PLUS and **`R`**. This family of statistical-computing environments featured vastly improved dynamic visualization capabilities, which allowed statisticians to identify outliers, trends and patterns in data that merited further study.



```{r, out.width = "300px"}
path <- paste0(getwd(), "/images/Tukey-EDA.jpg")
knitr::include_graphics(path)
```


## From S to R {.columns-2 .smaller}

**1975** Richard Becker and John Chambers, both statisticians working on data analysis at Bell Labs and both computer scientists, were frustrated with having to interrupt their data analysis to do the required programming. 

They wanted a more interactive process, where they could just type an expression at the command line and get back an immediate response, or better yet, a bar graph or other visual.

**1977: _Analalysis that was once difficult became easy in S_.**

**1991** R, based entirely on S, is created by Ross Ihaka and Robert Gentleman, borrowing all the hard-won lessons developed by the S community.

[Source: From S to R: 35 Years of Leadership in Statistical Computing](http://www.research.att.com/articles/featured_stories/2013_09/201309_SandR.html)

```{r, out.width = "150px"}
path <- paste0(getwd(), "/images/201309_RandS_timeline.jpg")
knitr::include_graphics(path)
```


